Structured Versus Decorated Cospans

### 2021-06-17 cospans_new.tex
working on the proof of Thm. 2.1

One can also check that $\Phi$ is an opfibration. Given a morphism $f \maps a \to a'$ in $\A$ and an object $(a \xrightarrow{i}b, x) \in \A/U$ over $a$, we shall show that a cocartesian lifting of $f$ is given by 
\[ (f,g,1) \maps (a \xrightarrow{i} b, x) \, \to \, (a' \xrightarrow{i'} b+_a a', F(g)(x)). \]
where $b +_a a'$, $i'$ and $g$ arise from the following pushout square in $\A$:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a$};
\node (B) at (1.5,0) {$a'$};
\node (C) at (0,-1) {$b$};
\node (D) at (1.5,-1) {$b+_a a'$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$f$} (B)
(A) edge node[left]{$i$} (C)
(B) edge [dashed] node[right]{$i'$} (D)
(C) edge [dashed] node[below]{$g$} (D);
\end{tikzpicture}
\]
To show that $(f,g,1)$ is an cocartesian lifting of $f$, suppose we are given a morphism 
\[    (f'',g'',h'') \maps (a \xrightarrow{i} b, x) \; \to \; (a'' \xrightarrow{i''} b'', x'')   \]
in $\A/U$ and a morphism $k \maps a \to a''$ in $\A$ such that $\Phi(f'',g'',h'') = k \circ f$.   To
show $\Phi$ is an opfibration, we need to show there exists a unique morphism
\[     (\tilde{f}, \tilde{g}, \tilde{h}) \maps (a' \xrightarrow{i'} b +_a a', F(g)(x)) \; \to \; (a'' \xrightarrow{i''} b'', x'') \]
such that $\Phi(\tilde{f}, \tilde{g}, \tilde{h}) = k$ and 
\[       (\tilde{f}, \tilde{g}, \tilde{h}) \circ (f,g,1) = (f'',g'',h'') .\]
To achieve this set $\tilde{f} = k$, $\tilde{h} = h''$, and define $\tilde{g}$ using
the universal property of the pushout:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a$};
\node (B) at (1.5,0) {$a'$};
\node (C) at (0,-1) {$b$};
\node (D) at (1.5,-1) {$b+_a a'$};
\node (E) at (2.5,-1.5) {$b''$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$f$} (B)
(A) edge node[left]{$i$} (C)
(B) edge node[left]{$i'$} (D)
(C) edge node[above]{$g$} (D)
(C) edge [bend right] node[below]{$g''$} (E)
(B) edge [bend left] node[right]{$i'' k$} (E)
(D) edge [dashed] node[left]{$\exists! \, \tilde{g} \phantom{a}$} (E);
\end{tikzpicture}
\]
after noting that $g'' i = i' f'' = i'' k f $ by the commuting square condition for the morphisms
in $\A/U$.   One can check that with these choices, $(\tilde{f},\tilde{g}, \tilde{h})$ obeys the desired conditions and is the unique morphism to do so.

@@

Denoting the cocartesian lifting of a morphism $f$ by $\textnormal{cocart}(f)$, we have that $\textnormal{cocart}(f_1) \otimes \textnormal{cocart}(f_2) \cong \textnormal{cocart}(f_1+f_2)$ by the universal map between two colimits of the same diagram: 
\[  (U(x_1)+_{a_1}b_1)+(U(x_2)+_{a_2}b_2) \cong U(x_1 \otimes x_2)+_{a_1+a_2}(b_1+b_2). \]

Lastly, the monoidal bifibration $\Phi$ satisfies the Beck--Chevalley condition. Given a pushout square in $\A$:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a$};
\node (B) at (1.5,0) {$b$};
\node (C) at (0,-1) {$c$};
\node (D) at (1.5,-1) {$d$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$h$} (B)
(A) edge node[left]{$k$} (C)
(B) edge node[right]{$g$} (D)
(C) edge node[below]{$f$} (D);
\end{tikzpicture}
\]
we need to show that the following square commutes up to a natural isomorphism:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$(\A/U)_a$};
\node (B) at (1.5,0) {$(\A/U)_b$};
\node (C) at (0,-1) {$(\A/U)_c$};
\node (D) at (1.5,-1) {$(\A/U)_d$};
\node (E) at (0.75,-0.5) {$\Searrow \theta$};
\path[->,font=\scriptsize,>=angle 90]
(B) edge node[above]{$h^*$} (A)
(A) edge node[left]{$k_!$} (C)
(B) edge node[right]{$g_!$} (D)
(D) edge node[below]{$f^*$} (C);
\end{tikzpicture}
\]
Let $(b,x,b \xrightarrow{i} U(x))$ be an object of $(\A/U)_b$ where $x=(b',p) \in \A \times F(b')$ so that $U(x)=b'$. Going left, we precompose with $h \maps a \to b$ to obtain the object $(a,x,a \xrightarrow{i \circ h} U(x))$ of $(\A/U)_a$. Then going down, we push forward along $k$:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a$};
\node (B) at (1.5,0) {$U(x)$};
\node (C) at (0,-1) {$c$};
\node (D) at (1.5,-1) {$c+_{a}U(x)$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i \circ h$} (B)
(A) edge node[left]{$k$} (C)
(B) edge node[right]{$$} (D)
(C) edge node[below]{$\psi$} (D);
\end{tikzpicture}
\]
to obtain the object $(c,z,c \xrightarrow{\psi} U(z))$ of $(\A/U)_c$, where $z \in \A \times F(c+_a U(x))$ is given by $z=(c+_a U(x),q)$ where $q$ is given by
\[ 1 \xrightarrow{p} F(U(x)) \xrightarrow{F(\zeta)} F(0+_0 U(x)) \xrightarrow{F(!_c +_{!_a} 1)} F(c+_a U(x)). \]
%and so by construction, we have that 
%\[  U(z)=U(\bot_c+_{\bot_a}x)=c+_a U(x).\]
%{\john Need to refer to a proposition where we show $U$ preserves pushouts strictly.}
 For the other route, first going down, we push the object $(b,x,b \xrightarrow{i} U(x))$ forward along $g \maps b \to d$ by taking the following pushout:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$b$};
\node (B) at (1.5,0) {$U(x)$};
\node (C) at (0,-1) {$d$};
\node (D) at (1.5,-1) {$d+_{b}U(x)$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i$} (B)
(A) edge node[left]{$g$} (C)
(B) edge node[right]{$$} (D)
(C) edge node[below]{$\theta$} (D);
\end{tikzpicture}
\]
which yields the object $(d,z',d \xrightarrow{\theta} U(z'))$ of $(\A/U)_d$ where $z' \in \A \times F(d+_b U(x))$ is given by ${z'=(d+_b U(x),q')}$ where $q'$ is given by 
\[ 1 \xrightarrow{p} F(U(x)) \xrightarrow{F(\zeta)} F(0+_0 U(x)) \xrightarrow{F(!_d +_{!_b} 1)} F(d+_b U(x)). \]
We then precompose with $f \maps c \to d$ to obtain the object $(c,z',c \xrightarrow{\theta \circ f} U(z'))$ of $(\A/U)_c$.

@@

\end{tikzpicture}
\]
and consequently a natural isomorphism between our two objects of $(\A/U)_c$, showing that $\Phi$ satisfies the Beck--Chevalley condition.   We then define $F\lCsp$ to be $\Fr(\Phi)$.
\end{proof}

>>>>

One can also check that $\Phi$ is an opfibration. Given a morphism $f \maps a \to a'$ in $\A$ and an object $(a \xrightarrow{i}b, x) \in \A/U$ over $a$, we claim that a cocartesian lifting of $f$ is given by 
\[ (f,g,1) \maps (a \xrightarrow{i} b, x) \, \to \, (a' \xrightarrow{i'} b+_a a', F(g)(x)). \]
where $g, i'$ and $b +_a a'$ arise from this pushout square in $\A$:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a$};
\node (B) at (1.5,0) {$a'$};
\node (C) at (0,-1) {$b$};
\node (D) at (1.5,-1) {$b+_a a'$.};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$f$} (B)
(A) edge node[left]{$i$} (C)
(B) edge [dashed] node[right]{$i'$} (D)
(C) edge [dashed] node[below]{$g$} (D);
\end{tikzpicture}
\]
Suppose we are given a morphism 
\[    (f'',g'',h'') \maps (a \xrightarrow{i} b, x) \; \to \; (a'' \xrightarrow{i''} b'', x'')   \]
in $\A/U$ and a morphism $k \maps a \to a''$ in $\A$ such that $\Phi(f'',g'',h'') = k \circ f$.   To show that $(f,g,1)$ is a cocartesian lifting of $f$, we need to show there exists a unique morphism
\[     (\tilde{f}, \tilde{g}, \tilde{h}) \maps (a' \xrightarrow{i'} b +_a a', F(g)(x)) \; \to \; (a'' \xrightarrow{i''} b'', x'') \]
such that $\Phi(\tilde{f}, \tilde{g}, \tilde{h}) = k$ and 
\[       (\tilde{f}, \tilde{g}, \tilde{h}) \circ (f,g,1) = (f'',g'',h'') .\]
To achieve this we choose $\tilde{f} = k$, $\tilde{h} = h''$, and define $\tilde{g}$ using
the universal property of the pushout:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a$};
\node (B) at (1.5,0) {$a'$};
\node (C) at (0,-1) {$b$};
\node (D) at (1.5,-1) {$b+_a a'$};
\node (E) at (2.5,-1.5) {$b''$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$f$} (B)
(A) edge node[left]{$i$} (C)
(B) edge node[left]{$i'$} (D)
(C) edge node[above]{$g$} (D)
(C) edge [bend right] node[below]{$g''$} (E)
(B) edge [bend left] node[right]{$i'' k$} (E)
(D) edge [dashed] node[below left]{$\exists! \, \tilde{g} \phantom{a}$} (E);
\end{tikzpicture}
\]
after noting that $g'' i = i' f'' = i'' k f $ by the commuting square condition for the morphisms
in $\A/U$.   One can check that with these choices, $(\tilde{f},\tilde{g}, \tilde{h})$ obeys the desired conditions and is the unique morphism to do so.

It follows that $\Phi$ is a bifibration.  Denoting the cocartesian lifting of a morphism $f$ by $\textnormal{cocart}(f)$, we have that $\textnormal{cocart}(f_1) \otimes \textnormal{cocart}(f_2) \cong \textnormal{cocart}(f_1+f_2)$ by the universal map between two colimits of the same diagram.  Thus, $\Phi$ is a monoidal bifibration.

@@

Lastly, we need to check that $\Phi$ satisfies the Beck--Chevalley condition. Given a pushout square in $\A$:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a$};
\node (B) at (1.5,0) {$b$};
\node (C) at (0,-1) {$c$};
\node (D) at (1.5,-1) {$d$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$h$} (B)
(A) edge node[left]{$k$} (C)
(B) edge node[right]{$g$} (D)
(C) edge node[below]{$f$} (D);
\end{tikzpicture}
\]
we need to show that the natural transformation $\theta$, defined as in \cref{subsec:fibrations}, is an isomorphism:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$(\A/U)_a$};
\node (B) at (1.5,0) {$(\A/U)_b$};
\node (C) at (0,-1) {$(\A/U)_c$};
\node (D) at (1.5,-1) {$(\A/U)_d$.};
\node (E) at (0.75,-0.5) {$\Searrow \theta$};
\path[->,font=\scriptsize,>=angle 90]
(B) edge node[above]{$h^*$} (A)
(A) edge node[left]{$k_!$} (C)
(B) edge node[right]{$g_!$} (D)
(D) edge node[below]{$f^*$} (C);
\end{tikzpicture}
\]
Let $(b \xrightarrow{i} e, x)$ be an object of $(\A/U)_b$. Going left, we precompose with $h \maps a \to b$ to obtain the object $(a \xrightarrow{i \circ h} e, z)$ of $(\A/U)_a$. Then going down, we push forward along $k$:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a$};
\node (B) at (1.5,0) {$e$};
\node (C) at (0,-1) {$c$};
\node (D) at (1.5,-1) {$c+_{a}e$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i \circ h$} (B)
(A) edge node[left]{$k$} (C)
(B) edge node[right]{$$} (D)
(C) edge node[below]{$\psi$} (D);
\end{tikzpicture}
\]
to obtain the object $(c \xrightarrow{\psi} c+_a e, F(\psi)(x))$ of $(\A/U)_c$.
%and so by construction, we have that 
%\[  U(z)=U(\bot_c+_{\bot_a}x)=c+_a U(x).\]
%{\john Need to refer to a proposition where we show $U$ preserves pushouts strictly.}
 For the other route, first going down, we push the object $(b,x,b \xrightarrow{i} U(x))$ forward along $g \maps b \to d$ by taking the following pushout:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$b$};
\node (B) at (1.5,0) {$U(x)$};
\node (C) at (0,-1) {$d$};
\node (D) at (1.5,-1) {$d+_{b}U(x)$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i$} (B)
(A) edge node[left]{$g$} (C)
(B) edge node[right]{$$} (D)
(C) edge node[below]{$\theta$} (D);
\end{tikzpicture}
\]
which yields the object $(d,z',d \xrightarrow{\theta} U(z'))$ of $(\A/U)_d$ where $z' \in \A \times F(d+_b U(x))$ is given by ${z'=(d+_b U(x),q')}$ where $q'$ is given by 
\[ 1 \xrightarrow{p} F(U(x)) \xrightarrow{F(\zeta)} F(0+_0 U(x)) \xrightarrow{F(!_d +_{!_b} 1)} F(d+_b U(x)). \]
We then precompose with $f \maps c \to d$ to obtain the object $(c,z',c \xrightarrow{\theta \circ f} U(z'))$ of $(\A/U)_c$.

@@

\end{tikzpicture}
\]
and consequently a natural isomorphism between our two objects of $(\A/U)_c$, showing that $\Phi$ satisfies the Beck--Chevalley condition.

Having checked the hypotheses of \cref{lem:shulman}, we then define $F\lCsp$ to be $\Fr(\Phi)$.
\end{proof}
